---
title: "Advanced uses of DiffSegR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Advanced uses of DiffSegR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r,  include = FALSE}
working_directory <- "./DIFFSEGR_TEST"
dir.create(working_directory, showWarnings = FALSE)
sample_info <- data.frame(
  sample    = c("pnp1_1_1", "pnp1_1_2", "wt_1", "wt_2"),
  condition = rep(c("pnp1_1", "wt"), each = 2),
  replicate = rep(1:2,2),
  bam       = sapply(
    c("pnp1_1_1_ChrC_71950_78500.bam", 
      "pnp1_1_2_ChrC_71950_78500.bam",
      "wt_1_ChrC_71950_78500.bam",
      "wt_2_ChrC_71950_78500.bam"
    ),
    function(bam) system.file("extdata", bam, package = "DiffSegR")
  ),
  coverage  = file.path(
    working_directory,
    paste0(c("pnp1_1_1", "pnp1_1_2", "wt_1", "wt_2"), ".rds")
  ),
  isPairedEnd    = rep(FALSE, 4),
  strandSpecific = rep(1, 4)
)
library(DiffSegR)
nb_threads_tot = 4
nb_threads_locus = 4
data <- newExperiment(
  sampleInfo   = sample_info,
  loci         = data.frame(
    seqid      = "ChrC", 
    chromStart = 71950, 
    chromEnd   = 78500,
    locusID    = "psbB_petD"
  ),
  referenceCondition = "wt",
  otherCondition     = "pnp1_1",
  nbThreads          = nb_threads_tot,
  nbThreadsByLocus   = nb_threads_locus
)
coverage(
  data         = data,
  coverageType = "average"
)
```

Before using this tutorial, we recommend that you read through our [introductory tutorial](introductory_tutorial.html).


# Overview

In this tutorial we explore more advanced uses of `DiffSegR`:

 1. How to explore a range of segmentation models ?
 2. How to select a subset of segments whose absolute $\log_2$-FC passes a threshold while keeping statistical guarantees on the true discovery proportion in this subset ? (Post hoc inference) 

# Exploring more segmentations  

After visual inspection, one can disagree with the segmentation of the $\log_2$-FC
proposed by **DiffSegR**. It can happen when spurious changepoints, resulting 
from an over-segmentation of the $\log_2$-FC, seems to be added. In this 
case it is often useful to explore several different segmentations in 
order to choose a more suitable hyperparameter $\alpha$ value.

To explore all reachable segmentations between 1 to `Kmax` segments, one can pass
`segmentNeighborhood = TRUE` and specify the value of `Kmax` to `segmentationLFC()`.
This step can be time consuming. A grid search approach is also implemented in
**DiffSegR** (see `gridSearch` and `alphas` parameters of `segmentationLFC()`).

```{r}
features <- segmentationLFC(
  data                = data, 
  outputDirectory     = working_directory,
  segmentNeighborhood = TRUE,
  Kmax                = 50
)
```

The figures showing the path of all reachable segmentations up to `Kmax` segments of
the log2-FC on reverse and forward strands are accessible in the working 
directory under the name *all_models_minus.rds* and *all_models_plus.rds*. 

A good practice is to select a robust segmentation. Visually it is equivalent 
to pick the elbow of this curve. In this example we can select 
$\alpha=5.46$ symbolized by the red line.

```{r echo=FALSE, dpi=300, out.width='100%', fig.align='center', fig.cap="Path of all reachable segmentations up to 50 segments of the log2-FC on reverse strand. Corresponding alpha values on the log scale can be read on X-axis."}
all_models <- readRDS("DIFFSEGR_TEST/psbB_petD/all_models_minus.rds")
ggplot2::ggplot(
  all_models,
  ggplot2::aes(
    x = log2(alphas), 
    y = K
  )
)+
ggplot2::geom_line() +
ggplot2::geom_point(size=1) +
ggplot2::geom_vline(
  xintercept = log2(5.46), 
  color      = "red"
)+
ggplot2::geom_hline(
  yintercept = 11, 
  color      = "blue"
)+
ggplot2::xlab("log2(alpha)")+
ggplot2::ylab("number of segments (K)")+
ggplot2::theme(
  text = ggplot2::element_text(size = 20)
) +
ggplot2::theme_bw()
```

# Post hoc inference

If the result of the DEA does not correspond to what the user expected,
they may tend to snoop in the data and eventually select a subset 
of segments of interest. For instance it is a common usage to define a *threshold* 
on the absolute $\log_2$-FC per-segment. In this case DiffSegR will declare 
differentially expressed the largest set of segments with an absolute 
$\log_2$-FC > *threshold* and a true discovery proportion lower bound choose by 
the user (``postHoc_tdpLowerBound``). This post-hoc 
lower bound is obtained by controlling the joint error rate (JER) using the 
Simes family of thresholds (Blanchard et al. 2020).


To use post hoc inference one have to specify :

  1. its decision rule, i.e a predicate function that returns a single `TRUE` 
  or ``FALSE`` if the segment on which it is applied meets the conditions defined in 
  the predicate. The criteria used in the predicate have to be defined for each
  segment in ``SummarizedExperiment::mcols(dds)``. Here, ``predicate = function(x) 2^abs(x$log2FoldChange)>2`` ;
  2. the minimum true discovery proportion among the subset of segments declared 
  differentially expressed. Here, ``postHoc_tdpLowerBound = 0.05`` ;
  3. the significance level of the test procedure. Here, 
  ``postHoc_significanceLevel = 0.05``.

```{r,  include = FALSE}
SExp <- counting(
  data     = data, 
  features = features
)
```

```{r}
dds <- dea(
  SExp                      = SExp,
  design                    = ~ condition,
  sizeFactors               = rep(1,4),
  predicate                 = function(x) 2^abs(x$log2FoldChange)>2,
  postHoc_significanceLevel = 0.05,
  postHoc_tdpLowerBound     = 0.95
)

#- check fold-change of differentially expressed segments ---------------------#
DERs <- dds[SummarizedExperiment::mcols(dds)$DER,]
2^abs(SummarizedExperiment::mcols(DERs)$log2FoldChange)
```

# References 

Blanchard, G., Neuvial, P. and Roquain, E. Post hoc confidence bounds on false 
positives using reference families. Ann. Statist 48(3), 1281-1303 (2020) [doi.org/10.1214/19-AOS1847](https://hal.science/hal-01483585v5/document).
