---
title: "Multi-loci analysis"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Multi-loci analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Before using this tutorial, we recommend that you read through our [introductory tutorial](introductory_tutorial.html).

DiffSegR can analyze multiple loci of interest, including across various 
chromosomes. To do so you can add one line per locus to the ``data.frame`` accepted as 
input by `newExperiment()` (see hereafter). Alternatively, the function also accepts a 
path to a tabular format file with similar columns (`seqid`, `chromStart`, 
`chromEnd`, `locusID`). The rest of the analysis can be conducted similarly to 
the analysis of a single locus.

```{r, eval = FALSE}
data <- newExperiment(
  sampleInfo   = sample_info,
  loci         = data.frame(
    seqid      = c("ChrC", "Chr1"), 
    chromStart = c(71950, 2421089), 
    chromEnd   = c(78500, 2422066),
    locusID    = c("psbB_petD", "F24B9")
  ),
  referenceCondition = "wt",
  otherCondition     = "pnp1_1",
  nbThreads          = 4,
  nbThreadsByLocus   = 2,
  coverage           = working_directory
)
```

Note : Here, the total number of threads used for the analysis is 4, with 2 
allocated per locus. Since the analysis involves two loci, at the appropriate 
time, both will be loaded into memory and processed simultaneously.